Temporal Evolution of Flow in Pore-Networks:\\ From Homogenization to Instability

### 2022-03-06 old_files/june14/supp.tex
Updates from Overleaf

This condition guarantees that at each step a small amount of material is eroded and there is no sudden change in the network. We further test the convergence by decreasing $\max(\Delta r_{ij})$ to half and we observe that the average relative change in the flux vector is $\approx 1.2\%$, and the PDFs remain intact without any notable change. The network size for the 2d/3d networks is $N_x=100,N_y=50$/$N_x=50,N_y=12, N_z=12$ unless mentioned otherwise. In either case, the network has an order of $10^3$ edges. The code is publicly available in a GitHub repository \cite{githubrepo}.

@@

  \includegraphics[width=.75\textwidth]{FigS1.pdf} % grid-analytica

@@

Taking the Laplace transform of the above equation and defining $\tilde p(s)  \equiv \int_0^{\infty} p(q) e^{-qs} dq$ one obtains

@@

\newpage
\section{Robustness to topology and initial condition}
% 
In order to check the robustness of our result with respect to the topology of the network, we run our simulations on a three-dimensional random network with Voronoi cell initialization of nodes in space (Fig. \ref{SIfig:fig2-3d}), and also for a two dimensional topologically ordered network with a diamond grid (Fig. \ref{fig:fig2-diamond}), where  both networks are initialized with a uniform random distribution for the diameter of the tubes. We find that regardless of the topology of the network, an erosion dynamics with $n>3$ results in a homogenized network, while $n<3$ results in the channelization instability. We further check the effect of initial randomness on the fate of the network. We use a two-dimensional diamond-grid network with a narrow uniform initial distribution of tube diameters around an average diameter $d_0$ with only a very small variation ($3\%$), i.e., the tube diameters are sampled from $\mathcal{U}(d_0(1-\epsilon), d_0(1+\epsilon))$. We again find that for the networks with $n<3$ channels are formed, while for $n>3$ the network stays homogenized.  
%
\begin{figure}[htp]
     % \centering
      \includegraphics[width = 1.0\textwidth]{FigS2.pdf}
     \caption{Erosion in a topologically random 3D network of tubes with $N_x=20, N_y=12, N_z=12$ points for Voronoi cell initialization and an initial uniform broad distribution of tube diameters randomly sampled from $\mathcal{U}(1,14)$. Snapshots of the network, PDF of normalized fluid flux $q/\langle q \rangle$, and normalized edge radius distribution $r/\langle r \rangle$ at the initial time $t=0$, and also after $N$ erosion steps for different powers of erosion $n$ are shown. We stop the erosion after $N$ steps such that $\langle r\rangle=2r_0$ where $r_0 = \langle r_{t=0}\rangle$. The erosion law is based on Eq. (1) in the main text where different powers of $n$ correspond to different models of erosion.}\label{SIfig:fig2-3d}
 \end{figure}
%
\begin{figure}[htp]
    % \centering
 \includegraphics[width=0.75\textwidth]{FigS3.pdf} % Fig2-Sdiamond.pdf
     \caption{Erosion in a diamond grid network with $N_x=100, N_y=50$ randomly distributed nodes and an initially broad  distribution of tube diameter randomly sampled from $\mathcal{U}(1,14)$. The initial condition is shown with the label $t=0$ in the first row. Each row afterward corresponds to the simulation result after $N$ steps such that $\langle r\rangle=2r_0$ where $r_0 = \langle r_{t=0}\rangle$ or twice the initial average radius. The erosion law is based on Eq. (1) in the main text where different powers of $n$ correspond to different models of erosion. The first column is a snapshot of the pore network, the second column is the PDF of normalized fluid flux $q/\langle q \rangle$, and the last column is the PDF of normalized radius $r/\langle r\rangle$.}\label{fig:fig2-diamond}% 
\end{figure}
%
\begin{figure}[htp]
    % \centering
 \includegraphics[width=0.72\textwidth]{FigS4.pdf}
     \caption{Erosion in a structured diamond grid network with $N_x=50, N_y=50$ and an initially narrow distribution of tube diameters randomly sampled from  $\mathcal{U}\left(d_0(1-\epsilon), d_u(1+\epsilon)\right) $ where  $d_0$ is the average diameter and $\epsilon=0.03$. The initial condition is shown with the label $t=0$ in the first row. Each row afterward corresponds to the simulation result after $N$ steps such that $\langle r_{t=N}\rangle=2r_0$ where $r_0 = \langle r_{t=0}\rangle$ or twice the initial average radius. The erosion law is based on Eq. (1) in the main text where different powers of $n$ correspond to different models of erosion. The first column is a snapshot of the pore network, the second column is the PDF of normalized fluid flux $q/\langle q \rangle$, and the last column is the PDF of normalized radius $r/\langle r\rangle$.}\label{fig:fig2}% 
\end{figure}

\newpage 
\newpage 
\newpage
\section{Average Change in the  Order parameter}
%
In order to quantify the network behavior shown in Fig. 4, we calculate the change in the order parameter for different $m,n$ averaged over 100 simulations with different random initial conditions, and the heat-map results are shown in Fig.~\ref{fig:fig4_SI-2}. The positive or negative change in the order parameter shows the network's change toward homogenization or channelization. The boundary between the two phases (homogenization and channelization) calculated using the simple model introduced in the main text is shown with a solid black line here, and it can be seen that it agrees well with the order parameter change. 
%
\begin{figure}[htp]
    % \centering
    \includegraphics[width = 0.70\textwidth]{FigS5.pdf}
    \caption{The heat map for the average change in the order parameter corresponding to the networks shown in Fig. 4 of the main text. The order parameter shown in the heat map here is an average of 100 different simulations with different random initialization of tube diameters. The black line shows the boundary between two phases calculated using the model introduced in the main text.}\label{fig:fig4_SI-2}
\end{figure}

@@

% \clearpage
\bibliography{ref}% Produces the bibliography via BibTeX.

\end{document}
%

%###############################################################
%###############################################################
%###############################################################
%###############################################################
%###############################################################
%###############################################################
%###############################################################

\newpage
\section{Clogging behavior at initial times}
% 
Considering a small change in the radius of edges, and as a result, the conductance of the edges $\mathbf{C}+\delta \mathbf{C}$, using Eq. \eqref{main-solve-eq} we find that 
%
\begin{align}
    \ket{P'} &=  \ket{P} -  \left( \Delta_n^\top\mathbf{C} \Delta_n\right)^{-1}\left[  \left( \Delta_n^\top \delta \mathbf{C} \Delta_n\right)\ket{P} + \left( \Delta_n^\top\delta \mathbf{C} \Delta_b\right) \ket{P_{BC}}\right],
\end{align}
%
where we used the fact that for an invertible matrix $\mathbf{A}$ we have 
%
\begin{align}
    \left(\mathbf{A} + \delta \mathbf{A} \right)^{-1} = \mathbf{A}^{-1} - \left( \mathbf{A}^{-1} \delta \mathbf{A}\right) \mathbf{A}^{-1} + \left( \mathbf{A}^{-1}  \delta \mathbf{A}\right)^{2}\mathbf{A}^{-1}  + \cdots.
\end{align}
%
Calculating for the changes in the fluid flux, we find that 
%
\begin{align}
\ket{q'} & = \left( \mathbf{C}+ \delta \mathbf{C}  \right)\left( \Delta_b \ket{P_{BC}} +  \Delta_n \ket{P'} \right) \\
& = \ket{q} + \delta\mathbf{C} \mathbf{C}^{-1}\ket{q} - \mathbf{C} \Delta_n \left( \Delta_n^\top\mathbf{C} \Delta_n\right)^{-1}\left[  \left( \Delta_n^\top \delta \mathbf{C} \Delta_n\right)\ket{P} + \left( \Delta_n^\top\delta \mathbf{C} \Delta_b\right) \ket{P_{BC}}\right].
\end{align}
%
The above equation can then be used to study the evolution of order parameter in the networks. Note that during the blockage of pore-throats, the connectivity between the nodes changes and as a result network's connectivity matrix $\Delta$ changes and the above approximation will not hold anymore.

%###############################################################
%###############################################################
%###############################################################
%###############################################################
%###############################################################
%###############################################################
%###############################################################

% ****** End of file apssamp.tex ******

 If we set
$\eta(w) = \delta(w-1/2)$, then the flow at each node has probability of
$1/2$ to move to either of the neighboring sites. As a result the flow at a given depth $L$ will become homogeneous as $P_L(Q) = \delta (Q-\bar{Q})$ where $\bar{Q}$
is the average of input fluid flow. This results is the same as having the same
diameter for all the edges which results in a singular solution in the structured grid.

@@

\subsection*{$P(Q)$ decays faster than $Q^{-n}$ for any $n$}
We first show that $P(Q)$ decays faster than any power of $Q$ for any
weight distributions. We expand the Laplace transform using a Taylor expansion as 
  \tilde P(s) = 1 + \sum_j \tilde P_{j} s^{j}.
Inserting this expansion into Eq. \eqref{eq:main-laplace}, we obtain
  1 + \sum_j \tilde P_{j} s^{j}  & = \lp  \int_0^1 d w \eta(w) \lb 1 + \sum_j \tilde P_{j} (sw)^{j}\rb    \rp^{N}  = \lp  1 + \sum_j \tilde P_j \langle w^{j}\rangle  s^{j}   \rp^{N}
where $\langle w^j \rangle = \int_0^1 \eta(w) w^j dw$ is the $j$-th
moment of the weights $w$. In the above equation we can observe that
  P_j \lp N \langle w^j\rangle -1 \rp = G(P_{j-1}, P_{j-1}, \cdots, P_1)
In the above equation $P_j$ can only diverge only if $N\langle w^j
\rangle -1=0$. If $w$ can only be zero and one, then $\langle
w^j\rangle =\langle w \rangle = 1/N$ and the coefficient can become
zero! Other than this special distribution, when $0 \leq w \leq 1$, then
  & w\in [0,1] \rightarrow w>w^2>w^3> \cdots \\
  & \forall k>1 \rightarrow \langle w^k \rangle < \langle w \rangle = \frac{1}{N} 
which means that $N\langle w^j \rangle -1\neq 0$ and as a result
$P_j$ is finite. $P_j$ on the other hand is the $j$-th moment or
$\langle Q^j\rangle$ and it is finite. If $P\approx Q^{-n}$, then the
$n$-th moment will not be finite i.e. $\langle Q^{n+1}\rangle =
P_{n+1}$ is not finite. We can then conclude that $P(Q)$ should go to
zero faster than any $Q^{-n}$ as $Q\to \infty$! In other words $d \log
P(Q) / d \log Q \to -\infty$ as $Q\to \infty$.  
\subsection*{Exponential Decay}
The main recursive governing equation for a diamond grid is  
  \tilde P(s) = \lp  \int_0^1 d w \eta(w) \tilde P(s w)  \rp^{2} \label{eq:main-recur}
% In order to have an approximation for $\eta(w)$ we consider the following situation:
At each node there are two neighboring nodes where the flux is distributed by
$w_1$ and $w_2$ among them. Assuming that $w_1$ is uniformly distributed
between $0$ and $1$, then $w_2 = 1-w_1$ (due to  conservation of
mass $w_1 + w_2 = 1$). Now $\eta(w)$ can be obtained using 
  & \eta(w) = M \int_0^1 dw_1 \delta(1-w_1-w) =  M, \\
  \text{since }& \int_0^1 \eta(w) dw = 1 \to M =1 \\
  & \eta (w) = 1
Just for completeness, if there are three neighboring connections ($N=3$), then we
have
  \eta(w) & = M \int_0^1 dw_1 \int_0^1 dw_2 \delta(1-w_1-w_2 - w)\\
          & = M(1-w)  \to \int_0^1\eta(w) dw = 1 \to M = \frac{1}{2} \\
  & \eta(w) = \frac{1}{2} (1-w)
As a result Eq.~\eqref{eq:main-recur} simplifies to
  \tilde P(s) = \lp  \int_0^1 d w  \tilde P(s w)  \rp^{2} 
In order to solve the above equation, we assume $\tilde V (s) =
\sqrt{\tilde{P}(s)}$.
  \tilde{V}(s) & = \int_0^1 dw \tilde V (sw) = \int_0^s \frac{du}{s}  \tilde{V}^{2}(u) \\
  s\tilde{V}(s) & = \int_0^s du  \tilde{V}^2(u) \\
Taking the differentiation with respect to $s$ yields
  \tilde{V}(s) + s \frac{d \tilde{V}(s)}{d s} = \tilde{V}^2(s) \\
  \frac{d \tilde{V}}{\tilde{V}^2-\tilde{V}} = \frac{ds}{s}  \\
  \log\lp \frac{1-\tilde{V}}{\tilde{V}} \rp  = \log(s)\\
  \tilde{V}(s) = \frac{1}{1+Cs} 
We set the mean to $1$ i.e.,
  \int_0^{\infty} Q P(Q) dQ = 1 \to \frac{d\tilde{P}}{ds}|_{s=0}  = -1\\
  \frac{d \tilde{V}}{ds}  = \frac{1}{2 \sqrt{\tilde{P}(s)}} \frac{d \tilde{P}(s)}{d s}   \to \frac{d\tilde{V}(s)}{ds}|_{s=0} =\frac{-1}{2} \\
  \frac{d\tilde{V}(s)}{ds} = \frac{-C}{(1+Cs)^{2}}|_{s=0} = -C = \frac{-1}{2}  \to C=\frac{1}{2} 
As a result, we have
  \tilde{P}(s) = \lp\frac{1}{1+s/2} \rp^2 \\
  P(Q) = 4Q e^{-2Q}
We can then show that for a uniform distribution of $w$'s, the PDF of
$Q$'s become an exponential tail. For other distribution of $w$'s, we
similarly find a similar trend which can be seen in the numerical simulations. 
% Actually, we showed that the PDF of $Q$ should decay faster than any power law for any distribution of
$w$. 
Note that in the distribution with exponential tail, the only parameter that determines the the PDF of the average fluid flow. This shows that why we observe the universal behaviour for the PDF of normalized fluid flux where the average of normalized fluid flux is unity.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%5

For each resistor we know that
%
\begin{align}
    q_{e} = R_{e} \Delta P_{e}
\end{align}
%
Assume that matrix $\mathbf{C}$ relates the pressure at nodes to the pressure difference, where 
%
\begin{align}
    \ket{\Delta P_e} = \mathbf{C} \ket{P_n}, \quad \ket{q_e} = \mathbf{R} \ket{\Delta P_e}, \quad \ket{q_n} = \mathbf{C}^\top \ket{q_e} \\
    \ket{q_n} = \mathbf{C}^\top \mathbf{R} \mathbf{C}\ket{P_n}
\end{align}
%
\begin{align}
    \mathbf{C}^\top \mathbf{R} \mathbf{C}\begin{bmatrix} P_1 \\ P_2 \\ \vdots \\ P_n
    \end{bmatrix}
    = \ket{q} \to \begin{bmatrix} \mathbf{C_{b}}^\top \mathbf{R} \mathbf{C_{b}}  & \vline &    \mathbf{C_{b}}^\top \mathbf{R} \mathbf{C_{n}} \\ \hline  
    \mathbf{C_{n}}^\top \mathbf{R} \mathbf{C_{b}} & \vline & \mathbf{C_{n}}^\top \mathbf{R} \mathbf{C_{n}}
    \end{bmatrix} \begin{bmatrix} P^{BC}_1 \\ P^{BC}_2 \\ \vdots \\ \hline  P_n \\ \vdots \\ P_{N_n}
    \end{bmatrix} = \begin{bmatrix} q_1 \\ q_2 \\ \vdots \\ \hline  0 \\ \vdots \\ 0\end{bmatrix} \\
    \begin{bmatrix} 
    \mathbf{A}_{bb} & \mathbf{A}_{bn} \\
    \mathbf{A}_{nb} & \mathbf{A}_{nn}       
    \end{bmatrix} \begin{bmatrix} P^{BC}_1 \\ P^{BC}_2 \\ \vdots \\ \hline  P_n \\ \vdots \\ P_{N_n}
    \end{bmatrix} = \begin{bmatrix} q_1 \\ q_2 \\ \vdots \\ \hline  0 \\ \vdots \\ 0\end{bmatrix} \to \begin{cases} 
        \mathbf{A}_{bb} \ket{P_{BC}} + \mathbf{A}_{bn}  \ket{P_{n}} = \ket{q} \\
    \mathbf{A}_{nb}\ket{P_{BC}} +  \mathbf{A}_{nn} \ket{P_n} = 0 
    \end{cases} \\
     \mathbf{A}_{nn} \ket{P_n} =- \mathbf{A}_{nb}\ket{P_{BC}} \to \mathbf{A}_{nn} \ket{P_n} = \mathbf{b}
\end{align}

In solving our problem, we have 
 \mathbf{C}_{n}^\top \mathbf{R} \mathbf{C}_{n}\ket{P} =     - \mathbf{C}_{n}^\top \mathbf{R} \mathbf{C}_{b}\ket{P}_{BC}  \to \mathbf{A}_{nn} \ket{P_n} = \mathbf{B}\ket{P}_{BC}
where $\mathbf{b}$ comes from the boundary conditions. When the pressures are found then $\ket{q} = \mathbf{R}\mathbf{C} \ket{P}$, we can obtain the new fluxes. Now the question is that if $\mathbf{R} = \mathbf{R}+\delta \mathbf{R}$, then what happens to the $\ket{q}$ with respect to the first order of perturbation? if we set $\mathbf{R} = \mathbf{R}+\delta \mathbf{R}$ then
\mathbf{C}_{n}^\top \left( \mathbf{R} + \delta \mathbf{R} \right) \mathbf{C}_{n}\ket{P'} &= - \mathbf{C}_{n}^\top \left( \mathbf{R} + \delta \mathbf{R} \right) \mathbf{C}_{b}\ket{P}_{BC} \\
(\mathbf{A} + \delta \mathbf{A}) \ket{P'} &= (\mathbf{B} + \delta \mathbf{B})\ket{P_{BC}} \\
\ket{P'} &=(\mathbf{A} + \delta \mathbf{A})^{-1} (\mathbf{B} + \delta \mathbf{B})\ket{P_{BC}}  \\
\ket{P'} &= \left( \mathbf{A}^{-1} - \left( \mathbf{A}^{-1} \delta \mathbf{A}\right) \mathbf{A}^{-1}\right) (\mathbf{B} + \delta \mathbf{B})\ket{P_{BC}} \\
\ket{P'} &= \mathbf{A}^{-1} \mathbf{B} \ket{P_{BC}} + \mathbf{A}^{-1}\delta \mathbf{B}\ket{P_{BC}} - \mathbf{A}^{-1} \delta \mathbf{A} \mathbf{A}^{-1} \mathbf{B} \ket{P_{BC}} \\
\ket{P'} &= \ket{P} + \mathbf{A}^{-1}\delta \mathbf{B}\ket{P_{BC}} - \mathbf{A}^{-1} \delta \mathbf{A} \ket{P}
where we used the fact that 
    \left(\mathbf{A} + \delta \mathbf{A} \right)^{-1} = \mathbf{A}^{-1} - \left( \mathbf{A}^{-1} \delta \mathbf{A}\right) \mathbf{A}^{-1} + \left( \mathbf{A}^{-1}  \delta \mathbf{A}\right)^{2}\mathbf{A}^{-1}  + \cdots
The fluid flow can then be obtained 
    \ket{q'} = \left( \mathbf{R}+ \delta \mathbf{R}\right) \mathbf{C} \ket{P'}    \\
    \text{where} \quad \mathbf{A} = \mathbf{C}^\top\mathbf{R} \mathbf{C},\quad \delta \mathbf{A} = \mathbf{C}^\top \delta \mathbf{R} \mathbf{C}
Upto the first order, we find that
    \ket{P'} &= \ket{P} + \mathbf{A}^{-1}\delta \mathbf{B}\ket{P_{BC}} - \mathbf{A}^{-1} \delta \mathbf{A} \ket{P} \\
    & = \left(\mathbf{I} - \mathbf{A}^{-1} \delta \mathbf{A}  \right) \ket{P} + \mathbf{A}^{-1}\delta \mathbf{B}\ket{P_{BC}} \\
    & = \left( \mathbf{I} - \left( \mathbf{C}_n^\top\mathbf{R} \mathbf{C}_n\right)^{-1} \left( \mathbf{C}_n^\top \delta \mathbf{R} \mathbf{C}_n\right)\right)\ket{P} -  \left( \mathbf{C}_n^\top\mathbf{R} \mathbf{C}_n\right)^{-1} \left( \mathbf{C}_n^\top\delta \mathbf{R} \mathbf{C}_b\right) \ket{P_{BC}} \\
    & = \ket{P} -  \left( \mathbf{C}_n^\top\mathbf{R} \mathbf{C}_n\right)^{-1}\left[  \left( \mathbf{C}_n^\top \delta \mathbf{R} \mathbf{C}_n\right)\ket{P} + \left( \mathbf{C}_n^\top\delta \mathbf{R} \mathbf{C}_b\right) \ket{P_{BC}}\right] 
Calculating the fluid flow, we obtain
\ket{q'} & = \left( \mathbf{R}+ \delta \mathbf{R}  \right)\left( \mathbf{C}_b \ket{P_{BC}} +  \mathbf{C}_n \ket{P'} \right) \\
& = \left( \mathbf{R}+ \delta \mathbf{R}  \right)\left( \mathbf{C}_b \ket{P_{BC}} + \mathbf{C}_n \left\{ \ket{P} -  \left( \mathbf{C}_n^\top\mathbf{R} \mathbf{C}_n\right)^{-1}\left[  \left( \mathbf{C}_n^\top \delta \mathbf{R} \mathbf{C}_n\right)\ket{P} + \left( \mathbf{C}_n^\top\delta \mathbf{R} \mathbf{C}_b\right) \ket{P_{BC}}\right] \right\} \right) \\
& = \mathbf{R}\left( \mathbf{C}_b \ket{P_{BC}} +  \mathbf{C}_n \ket{P} \right) + \delta \mathbf{R}\left( \mathbf{C}_b \ket{P_{BC}} +  \mathbf{C}_n \ket{P} \right) - \mathbf{R} \mathbf{C}_n \left( \mathbf{C}_n^\top\mathbf{R} \mathbf{C}_n\right)^{-1}\left[  \left( \mathbf{C}_n^\top \delta \mathbf{R} \mathbf{C}_n\right)\ket{P} + \left( \mathbf{C}_n^\top\delta \mathbf{R} \mathbf{C}_b\right) \ket{P_{BC}}\right]  \\

>>>>

This condition guarantees that at each step a small amount of material is eroded and there is no sudden change in the network. We further test the convergence by decreasing $\max(\Delta r_{ij})$ to half and we observe that the average relative change in the flux vector is $\aprrox 1.2\%$, and the PDFs remain intact without any notable change. The network size for the 2d/3d networks is $N_x=100,N_y=50$/$N_x=50,N_y=12, N_z=12$ unless mentioned otherwise. In either case, the network has an order of $10^3$ edges. The code is publicly available in a GitHub repository \cite{githubrepo}.

@@

  \includegraphics[width=.75\textwidth]{./FigS1.pdf} % grid-analytica

@@

Taking the Laplace transform of the above equation and defining $\tilde p(s) & \equiv \int_0^{\infty} p(q) e^{-qs} dq$ one obtains

@@

\newpage
\section{Robustness to topology and initial condition}
% 
In order to check the robustness of our result with respect to the topology of the network, we run our simulations on a three-dimensional random network with Voronoi cell initialization of nodes in space (Fig. \ref{SIfig:fig2-3d}), and also for a two dimensional topologically ordered network with a diamond grid (Fig. \ref{fig:fig2-diamond}), where  both networks are initialized with a uniform random distribution for the diameter of the tubes. We find that regardless of the topology of the network, an erosion dynamics with $n>3$ results in a homogenized network, while $n<3$ results in the channelization instability. We further check the effect of initial randomness on the fate of the network. We use a two-dimensional diamond-grid network with a narrow uniform initial distribution of tube diameters around an average diameter $d_0$ with only a very small variation ($3\%$), i.e., the tube diameters are sampled from $\mathcal{U}(d_0(1-\epsilon), d_0(1+\epsilon))$. We again find that for the networks with $n<3$ channels are formed, while for $n>3$ the network stays homogenized.  
%
\begin{figure}[htp]
     % \centering
      \includegraphics[width = 1.0\textwidth]{./FigS2.pdf}
     \caption{Erosion in a topologically random 3D network of tubes with $N_x=20, N_y=12, N_z=12$ points for Voronoi cell initialization and an initial uniform broad distribution of tube diameters randomly sampled from $\mathcal{U}(1,14)$. Snapshots of the network, PDF of normalized fluid flux $q/\langle q \rangle$, and normalized edge radius distribution $r/\langle r \rangle$ at the initial time $t=0$, and also after $N$ erosion steps for different powers of erosion $n$ are shown. We stop the erosion after $N$ steps such that $\langle r\rangle=2r_0$ where $r_0 = \langle r_{t=0}\rangle$. The erosion law is based on Eq. (1) in the main text where different powers of $n$ correspond to different models of erosion.}\label{SIfig:fig2-3d}
 \end{figure}
%
\begin{figure}[htp]
    % \centering
 \includegraphics[width=0.75\textwidth]{./FigS3.pdf} % Fig2-Sdiamond.pdf
     \caption{Erosion in a diamond grid network with $N_x=100, N_y=50$ randomly distributed nodes and an initially broad  distribution of tube diameter randomly sampled from $\mathcal{U}(1,14)$. The initial condition is shown with the label $t=0$ in the first row. Each row afterward corresponds to the simulation result after $N$ steps such that $\langle r\rangle=2r_0$ where $r_0 = \langle r_{t=0}\rangle$ or twice the initial average radius. The erosion law is based on Eq. (1) in the main text where different powers of $n$ correspond to different models of erosion. The first column is a snapshot of the pore network, the second column is the PDF of normalized fluid flux $q/\langle q \rangle$, and the last column is the PDF of normalized radius $r/\langle r\rangle$.}\label{fig:fig2-diamond}% 
\end{figure}
%
\begin{figure}[htp]
    % \centering
 \includegraphics[width=0.72\textwidth]{./FigS4.pdf}
     \caption{Erosion in a structured diamond grid network with $N_x=50, N_y=50$ and an initially narrow distribution of tube diameters randomly sampled from  $\mathcal{U}\left(d_0(1-\epsilon), d_u(1+\epsilon)\right) $ where  $d_0$ is the average diameter and $\epsilon=0.03$. The initial condition is shown with the label $t=0$ in the first row. Each row afterward corresponds to the simulation result after $N$ steps such that $\langle r_{t=N}\rangle=2r_0$ where $r_0 = \langle r_{t=0}\rangle$ or twice the initial average radius. The erosion law is based on Eq. (1) in the main text where different powers of $n$ correspond to different models of erosion. The first column is a snapshot of the pore network, the second column is the PDF of normalized fluid flux $q/\langle q \rangle$, and the last column is the PDF of normalized radius $r/\langle r\rangle$.}\label{fig:fig2}% 
\end{figure}

\newpage 
\newpage 
\newpage
\section{Average Change in the  Order parameter}
%
In order to quantify the network behavior shown in Fig. 4, we calculate the change in the order parameter for different $m,n$ averaged over 100 simulations with different random initial conditions, and the heat-map results are shown in Fig.~\ref{fig:fig4_SI-2}. The positive or negative change in the order parameter shows the network's change toward homogenization or channelization. The boundary between the two phases (homogenization and channelization) calculated using the simple model introduced in the main text is shown with a solid black line here, and it can be seen that it agrees well with the order parameter change. 
%
\begin{figure}[htp]
    % \centering
    \includegraphics[width = 0.70\textwidth]{./FigS5.pdf}
    \caption{The heat map for the average change in the order parameter corresponding to the networks shown in Fig. 4 of the main text. The order parameter shown in the heat map here is an average of 100 different simulations with different random initialization of tube diameters. The black line shows the boundary between two phases calculated using the model introduced in the main text.}\label{fig:fig4_SI-2}
\end{figure}

@@

% \clearpage
% \bibliography{ref}% Produces the bibliography via BibTeX.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%5
%apsrev4-2.bst 2019-01-14 (MD) hand-edited version of apsrev4-1.bst
%Control: key (0)
%Control: author (8) initials jnrlst
%Control: editor formatted (1) identically to author
%Control: production of article title (0) allowed
%Control: page (0) single
%Control: year (1) truncated
%Control: production of eprint (0) enabled
\begin{thebibliography}{5}%
\makeatletter
\providecommand \@ifxundefined [1]{%
 \@ifx{#1\undefined}
}%
\providecommand \@ifnum [1]{%
 \ifnum #1\expandafter \@firstoftwo
 \else \expandafter \@secondoftwo
 \fi
}%
\providecommand \@ifx [1]{%
 \ifx #1\expandafter \@firstoftwo
 \else \expandafter \@secondoftwo
 \fi
}%
\providecommand \natexlab [1]{#1}%
\providecommand \enquote  [1]{``#1''}%
\providecommand \bibnamefont  [1]{#1}%
\providecommand \bibfnamefont [1]{#1}%
\providecommand \citenamefont [1]{#1}%
\providecommand \href@noop [0]{\@secondoftwo}%
\providecommand \href [0]{\begingroup \@sanitize@url \@href}%
\providecommand \@href[1]{\@@startlink{#1}\@@href}%
\providecommand \@@href[1]{\endgroup#1\@@endlink}%
\providecommand \@sanitize@url [0]{\catcode `\\12\catcode `\$12\catcode
  `\&12\catcode `\#12\catcode `\^12\catcode `\_12\catcode `\%12\relax}%
\providecommand \@@startlink[1]{}%
\providecommand \@@endlink[0]{}%
\providecommand \url  [0]{\begingroup\@sanitize@url \@url }%
\providecommand \@url [1]{\endgroup\@href {#1}{\urlprefix }}%
\providecommand \urlprefix  [0]{URL }%
\providecommand \Eprint [0]{\href }%
\providecommand \doibase [0]{https://doi.org/}%
\providecommand \selectlanguage [0]{\@gobble}%
\providecommand \bibinfo  [0]{\@secondoftwo}%
\providecommand \bibfield  [0]{\@secondoftwo}%
\providecommand \translation [1]{[#1]}%
\providecommand \BibitemOpen [0]{}%
\providecommand \bibitemStop [0]{}%
\providecommand \bibitemNoStop [0]{.\EOS\space}%
\providecommand \EOS [0]{\spacefactor3000\relax}%
\providecommand \BibitemShut  [1]{\csname bibitem#1\endcsname}%
\let\auto@bib@innerbib\@empty
%</preamble>
\bibitem [{git(2021)}]{githubrepo}%
  \BibitemOpen
  \href@noop {} {\bibinfo {title} {Network instability}},\ \bibinfo
  {howpublished} {\url{https://github.com/ahmadzareei/networkInstability}}
  (\bibinfo {year} {2021})\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Liu}\ \emph {et~al.}(1995)\citenamefont {Liu},
  \citenamefont {Nagel}, \citenamefont {Schecter}, \citenamefont {Coppersmith},
  \citenamefont {Majumdar}, \citenamefont {Narayan},\ and\ \citenamefont
  {Witten}}]{liu1995force}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {C.-h.}\ \bibnamefont
  {Liu}}, \bibinfo {author} {\bibfnamefont {S.~R.}\ \bibnamefont {Nagel}},
  \bibinfo {author} {\bibfnamefont {D.}~\bibnamefont {Schecter}}, \bibinfo
  {author} {\bibfnamefont {S.}~\bibnamefont {Coppersmith}}, \bibinfo {author}
  {\bibfnamefont {S.}~\bibnamefont {Majumdar}}, \bibinfo {author}
  {\bibfnamefont {O.}~\bibnamefont {Narayan}},\ and\ \bibinfo {author}
  {\bibfnamefont {T.}~\bibnamefont {Witten}},\ }\bibfield  {title} {\bibinfo
  {title} {Force fluctuations in bead packs},\ }\href@noop {} {\bibfield
  {journal} {\bibinfo  {journal} {Science}\ }\textbf {\bibinfo {volume}
  {269}},\ \bibinfo {pages} {513} (\bibinfo {year} {1995})}\BibitemShut
  {NoStop}%
\bibitem [{\citenamefont {Coppersmith}\ \emph {et~al.}(1996)\citenamefont
  {Coppersmith}, \citenamefont {Liu}, \citenamefont {Majumdar}, \citenamefont
  {Narayan},\ and\ \citenamefont {Witten}}]{coppersmith1996model}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {S.}~\bibnamefont
  {Coppersmith}}, \bibinfo {author} {\bibfnamefont {C.-h.}\ \bibnamefont
  {Liu}}, \bibinfo {author} {\bibfnamefont {S.}~\bibnamefont {Majumdar}},
  \bibinfo {author} {\bibfnamefont {O.}~\bibnamefont {Narayan}},\ and\ \bibinfo
  {author} {\bibfnamefont {T.}~\bibnamefont {Witten}},\ }\bibfield  {title}
  {\bibinfo {title} {Model for force fluctuations in bead packs},\ }\href@noop
  {} {\bibfield  {journal} {\bibinfo  {journal} {Physical Review E}\ }\textbf
  {\bibinfo {volume} {53}},\ \bibinfo {pages} {4673} (\bibinfo {year}
  {1996})}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {Alim}\ \emph {et~al.}(2017)\citenamefont {Alim},
  \citenamefont {Parsa}, \citenamefont {Weitz},\ and\ \citenamefont
  {Brenner}}]{alim2017local}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {K.}~\bibnamefont
  {Alim}}, \bibinfo {author} {\bibfnamefont {S.}~\bibnamefont {Parsa}},
  \bibinfo {author} {\bibfnamefont {D.~A.}\ \bibnamefont {Weitz}},\ and\
  \bibinfo {author} {\bibfnamefont {M.~P.}\ \bibnamefont {Brenner}},\
  }\bibfield  {title} {\bibinfo {title} {Local pore size correlations determine
  flow distributions in porous media},\ }\href@noop {} {\bibfield  {journal}
  {\bibinfo  {journal} {Physical Review Letters}\ }\textbf {\bibinfo {volume}
  {119}},\ \bibinfo {pages} {144501} (\bibinfo {year} {2017})}\BibitemShut
  {NoStop}%
\bibitem [{\citenamefont {Parsa}\ \emph {et~al.}(2021)\citenamefont {Parsa},
  \citenamefont {Zareei}, \citenamefont {Santanach-Carreras}, \citenamefont
  {Morris}, \citenamefont {Amir}, \citenamefont {Xiao},\ and\ \citenamefont
  {Weitz}}]{shima2021}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {S.}~\bibnamefont
  {Parsa}}, \bibinfo {author} {\bibfnamefont {A.}~\bibnamefont {Zareei}},
  \bibinfo {author} {\bibfnamefont {E.}~\bibnamefont {Santanach-Carreras}},
  \bibinfo {author} {\bibfnamefont {E.}~\bibnamefont {Morris}}, \bibinfo
  {author} {\bibfnamefont {A.}~\bibnamefont {Amir}}, \bibinfo {author}
  {\bibfnamefont {L.}~\bibnamefont {Xiao}},\ and\ \bibinfo {author}
  {\bibfnamefont {D.~A.}\ \bibnamefont {Weitz}},\ }\bibfield  {title} {\bibinfo
  {title} {Unexpected scaling of interstitial velocities with permeability due
  to polymer retention in porous media},\ }\href@noop {} {\bibfield  {journal}
  {\bibinfo  {journal} {submitted}\ } (\bibinfo {year} {2021})}\BibitemShut
  {NoStop}%
\end{thebibliography}%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\end{document}

@@
